The Effects of the NBA COVID Bubble on the NBA Playoffs

### 2022-01-30 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Discussion of significance of 3% change in made 2's

There are also surveys on the factors contributing to home-court advantage.
@Carron1992 gave four main game location factors for home and away
teams, namely, the crowd factor, which is the impact of fans cheering; learning
factors, which is an advantage from home teams from playing at
a familiar venue; travel factors, the idea that away teams may face
fatigue and jet lag from traveling; and, rule factors, which says that home teams
may benefit from some advantages in rules and officiating. They acknowledge that
these factors would all be removed if games were played at a neutral site even
if one team was designated as “home team”. This study was reviewed a decade later
by @Carron2005. The 2005 review goes over the new findings from studies
about the significance of these four game location factors. Since 1992 they found
that results on these four factors are mixed. However, there is some evidence
supporting crowd and travel factors impact games across all major sports. There is less
evidence suggesting learning and rule factors impact across the various collegiate 
and professional sports, regardless, the NBA is not a league which has rules that
may favor the home team like batting last in the MLB. One interesting finding cited by @Carron2005 is that in the absence of crowds result in overall performance increases.

@@

>>>>

There are also surveys on the factors contributing to home-court advantage.
@Carron1992 gave four main game location factors for home and away
teams, namely, the crowd factor, which is the impact of fans cheering; learning
factors, which is an advantage from home teams from playing at
a familiar venue; travel factors, the idea that away teams may face
fatigue and jet lag from traveling; and, rule factors, which says that home teams
may benefit from some advantages in rules and officiating. They acknowledge that
these factors would all be removed if games were played at a neutral site even
if one team was designated as “home team”. This study was reviewed a decade later
by @Carron2005. The 2005 review goes over the new findings from studies
about the significance of these four game location factors. Since 1992 they found
that results on these four factors are mixed. However, there is some evidence
supporting crowd and travel factors impact games across all major sports. There is less
evidence suggesting learning and rule factors impact across the various collegiate 
and professional sports, regardless, the NBA is not a league which has rules that
may favor the home team like batting last in the MLB. One interesting finding cited
by @Carron2005 is that in the absence of crowds result in overall performance increases.
Lastly, as discussed in the introduction @Lopez find evidence that geographic
factors like altitude may influence and strengthen home court advantage for teams in
high altitude regions.

@@

To get a bit more detailed, we saw an estimated 3\% increase in away two point
shooting percentage. Likewise, NBA teams shot
an average of about 53 two point shots per game, in our data, leading to a three 
point increase in total scoring. Even this
small swing in scoring for away teams can make a big difference in outcome of
games. Around 1/3 of NBA games end by a decision of six points or less, those
are the games where the three points matter most. Suppose there are 30 seconds
left and the away team is losing. If the away team has the ball, only down three, 
they can take their time finding a good shot which can
still either be a two or three, given the score and remaining time. In the context 
of the NBA bubble, this crucial possession also has the benefit of the away team
getting to work in silence without the jeers from enemy fans. However, in the
same situation while being down six the offense must rush to get a quick shot
off, which probably has to be a three pointer. This is a much more desperate
circumstance, less likely to have a positive outcome and even in the positive
outcome case you're still going to be down three and likely have to foul the
opposing team after you score. 
Alternatively, if the away team is playing defense with 30 seconds left, only down 
three, you can play regular defense without fouling, get a stop, then draw up a play 
to score a three pointer with the remaining six seconds. However, in the same scenario 
down six, you have no choice, but to foul. The free throws will likely put you down
seven or eight points, but the alternative is let the opponent run the 24 second shot 
clock down wasting time. Also, the away team may even be in a situation where they're
leading by the three additional points, which for the same logic as explained above 
puts them in far greater control of the game. Clearly, the three point swing can 
drastically change probability of winning for a team down the stretch of a game. 
Since, home teams picked up an additional two points per game on average vs. five 
for away teams, the three additional points from two point shooting appear to be the 
driving factor of away teams closing the gap with home teams.
